---
title: "Covid91 vaccine study Final2023"
author: "Estrella Aguilar "
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---



```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, covid19 = "covid", OK = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
Males = filter(FinalData, sex =="M")
Females = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
Druggies = filter(FinalData, DrugUser =="yes")
```





# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Code block modification 

I made some changes to the block of code on line 14, I used Tay's example as an example for what changes to make. I separated each variable into its own line and changed its name to Males, Females, LGBTQ and DrugUser(because druggies was not found). I also changed some formatting to seperate the people who got covid and the people who did not get covid. This allowed for all 4 groups to pop up in the environment panel. 

## Males


## Barplot 

```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot shows the amount of males who got Covid after getting the vaccine and taking a placebo. 

The plot above shows that 30% of the people that took the vaccine covid-19, while 69% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took the placebo contracted covid-19 versus those who got the vaccine. The vaccine appears to be effective in protection against covid-19. The odds of not getting covid-19 while taking the vaccine is 2.25:1. The men that took the vaccine were 2.25 times less likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and percent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot. I used Taylor's example and changed the code "guys" to "Males" to match up with my data. 


  
### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
The plot and numerical results shows that 69% of Males got covid after taking the placebo and only 30% got covid after getting the vaccine. A significantly larger percentage of those who took the placebo  contracted covid-19 than those who got the vaccine. The vaccine appears to provide more protection against Covid-19.The odds of not getting covid-19 while taking the vaccine is 2.25:1. The Males that took the vaccine were 2.25 times less likely to get covid-19 than those taking the placebo.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```


The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 1.098e-13. This is an extremely small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option.

The odd ratio above illustrates a ratio of  2.242522. the Males(men) who received the vaccine were 2.24 times less to get covid-19 than those who received the placebo. 


```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine for guys(males). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting Males from getting covid-19.

The odds ratio is approximately how effective the vaccine is. The odds ratio for the Males is 69%. 





## Females
```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot shows the amount of Females who got Covid after getting the vaccine and taking a placebo. 

In this plot, contrary to the previous, it appears that more people who were only given the placebo contracted covid-19. Only approximately 40% of the individuals who were given the vaccine were reported to have contracted covid-19. That means around 60% of the people that got covid-19 were the individuals that only received the placebo.The odds of getting covid-19 while taking the vaccine is 1.5. The gals that took the vaccine were 1.5 times  less likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 



The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The plot and numerical results shows that 60% of feamles got covid after taking the placebo and only 40% of females got covid after getting the vaccine. A significantly larger percentage of those who took the placebo  contracted covid-19 than those who got the vaccine. The vaccine appears to provide more protection against Covid-19.The odds of getting covid-19 while taking the vaccine is 1.5 for Females. The Females that took the vaccine were 1.5 times less likely to get covid-19 than those taking the placebo.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalesfish <- fisher.test(table1)
Femalesfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 3e-04. This is an extremely tiny number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.


The Fishers's Exact Test illustrates that the odds ratio is 1.5.
Females(Women) who got the vaccine were about 1.5 times less likely to get covid-19 than women that only took the placebo.


```{r}
effect <- Femalesfish$p.value/(Femalesfish$p.value +1)
effectivenessFemales = Femalesfish$estimate/(Femalesfish$estimate +1)*100
print(effectivenessFemales)
```
```{r}
low=(Femalesfish$conf.int[1]/(Femalesfish$conf.int[1] +1))*100
hi=(Femalesfish$conf.int[2]/(Femalesfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`


Above is the effectiveness for the covid-19 vaccine for Females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we cn conclude that the vaccine is effective in protecting Femaales from getting covid-19.The odds ratio is approximately how effective the vaccine is. The odds ratio for the Females is 60%.




## LGBTQ

```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot shows the amount of LGBTQ people who got Covid after getting the vaccine and taking a placebo. 

In this barplot, it appears that a sniffling 85% of the individuals that took the vaccine got covid-19 as well as the 15% who received the placebo. The odds of getting covid-19 while taking the vaccine is 0.18. The individuals who identified as gay and took the vaccine were 0.18 times as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of LGBTQ individuals that got covid-19 after taking the placebo and taking the vaccine. 
The LGBTQ participants that took the vaccine and contracted covid-19 was at an approximately shocking 85%! Those participants that only received the placebo only contacted covid-19 at 15%. The vaccine id evidently completely ineffective in protecting the LGBTQ community against covid-19 and taking it is not advised. 
The odds of getting covid-19 while taking the vaccine is 0.18. The individuals who identified as LGBTQ and took the vaccine were 0.18 more likely to get covid-19 than those taking the placebo.
### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```

The p-value is 7e-08. This in an incredibly small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.18
LGBTQ individuals who got the vaccine were about 0.18 times more likely to get Covid-19 than LGBTQ that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is completely ineffective in preventing the chances of getting covid-19 in individuals who identify as gay.

```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```
```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`



Above is the effectiveness for the covid-19 for those who identified as LGBTQ.  If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting people who identify as LGBTQ from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the LGBTQ folk be seen above as 15. This is completely ineffective. The people who take the vaccine are not better off then those who don't receive it at all. The effectiveness confidence interval for the people who identified as LGBTQ is 7.6 to 26.57.


## Druggies
```{r}
dd2 <- Druggies %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered  Druggies to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

This barplot shows that around 88% of the people who answered yes to drug use and received the vaccine contracted covid-19. This means that around 12% of the druggies that only took the placebo ended up getting covid-19.The odds of getting covid-19 while taking the vaccine is 0.13. The druggies that took the vaccine were 0.13 more likely to get covid-19 than those taking the placebo.




There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 




The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.
### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
The percentage of drug users that took the vaccine and contacted covid-19 is 88% Those that were only given the placebo and got covid-19 was almost 12%. It is clear that the vaccine to prevent covid-19 is ineffective in protecting drug users against covid-19 and  so getting this vaccine is not recommended. The odds of getting covid-19 while taking the vaccine is 0.13 The drugusers that took the vaccine were 0.13 more likely to get covid-19 than those taking the placebo.



### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Druggiesfish <- fisher.test(table1)
Druggiesfish
```
  
  
  
  
The p-value is 3e-09. This in a small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.13
drug users who got the vaccine were about 0.13 more likely to get Covid-19 than drug users that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is  ineffective in preventing the chances of getting covid-19 in drugusers.


```{r}
effect <- Druggiesfish$p.value/(Druggiesfish$p.value +1)
effectivenessDruggies = Druggiesfish$estimate/(Druggiesfish$estimate +1)*100
print(effectivenessDruggies)
```
```{r}
low=(Druggiesfish$conf.int[1]/(Druggiesfish$conf.int[1] +1))*100
hi=(Druggiesfish$conf.int[2]/(Druggiesfish$conf.int[2] +1))*100
options(digits=2)
```

  
Above is the effectiveness for the covid-19 vaccine for Druggies. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is not effective in protecting Druggies from getting covid-19.The odds ratio is approximately how effective the vaccine is. The odds ratio for the Druggies is 11%.

## Effectiveness Confidence Interval `r low`  to `r hi`


Above, is the effectiveness for the covid-19 for the druggies. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is not effective in protecting drug users from getting covid-19. 

This vaccine does not work in protecting drug users from contracting covid-19.

The odds ratio is approximately how effective the vaccine is. The odds ratio for the drug users can be seen above as 11. The effectiveness confidence interval for the drug users is 4.6 to 21.82.




# Overall Results and Conclusions

The covid-19 vaccine was effective in protecting the males against the contraction of the covid-19 The covid-19 effectiveness was `r effectivenessMales` for men.The men were protected the most with this particular vaccine.


The covid-19 vaccine was highly successful in protecting women against covid-19.The covid-19 effectiveness was `r effectivenessFemales`.

The covid-19 vaccine was not effective in treating people who identified as LGBTQ. The individuals who identified as LGBTQ and took the vaccine were 0.18 times more likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenessLGBTQ` for LGBTQ individuals.


The Druggies that took the vaccine were 0.13 times more likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenessDruggies` for Druggies.

If you are a Druggie or identify as LGBTQ it is not recommended to receive the vaccine. However, if you are a male or female, the vaccine was reported to have been effective.
The null that there is no difference in taking the vaccine can be accepted to the categories including druggies and LGBTQ individuals. the alternative, that the vaccine works and is effective, can be accepted for the Females and Males.


